---
title: "home_prices"
author: "mcdevitt"
date: "14 avril 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())

library(ggplot2)
library(MASS)

home_dir <- "~/_smu/_src/home_prices/"
setwd(home_dir)
data_dir <- "./data"


```

## House Prices - à la Kaggle

Ask a home buyer to describe their dream house, and they probably won't begin with the height of the basement ceiling or the proximity to an east-west railroad. But this playground competition's dataset proves that much more influences price negotiations than the number of bedrooms or a white-picket fence.  
With 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa, this competition challenges you to predict the final price of each home.  
  
Deliverables:  
Your group is to turn in a paper that is no more that 7 pages long (without the appendix). Please put your code in the appendix, but any graphs and tables in the body of the paper.   
  
Sample Format  
Required deliverables in the complete report.  The format of your paper (headers, sections, etc) is flexible although should contain the following information:  
  
Introduction   
  
Data Description   
(Where did the data come from?  How big is it? How many observations?  Where can we find out more?  What are the specific variables that we need to know to understand with respect to your analysis?)  
  
Analysis Question 1:  
	Restatement of Problem   
  
	Specify the Model  
	   
	Checking Assumptions   
		Residual Plots   
		Influential point analysis (Cook’s D and Leverage)  
		Make sure and address each assumption.  
  
	Comparing Competing Models  
		adj R2    
		Interval CVPress    
	  
	Parameter Interpretation  
		Interpretation   
		Confidence Intervals   
  
Conclusion  
	A short summary of the analysis.  
	  
  
Analysis Question 2  
Restatement of Problem   
  
Model Selection  
		Type of Selection  
			Stepwise  
Forward  
Backward  
CUSTOM 		   
  
		Checking Assumptions   
			Residual Plots  
			Influential point analysis (Cook’s D and Leverage)  
			Make sure and address each assumption  
  
		Comparing Competing Models  
			Adj R2     
			Interval CVPress     
			Kaggle Score   
  
		Conclusion: A short summary of the analysis.    
	  
  
Appendix  
	Well commented SAS Code for Analysis 1 and 2   
  
  
Rubric:  
Presentation (30%):  
	Organized paper with title, headings, subheadings, etc.  
		Labeled plots, figures, tables and charts.  
		Every plot, figure, table and chart included is referenced in the paper and vice versa.   
		No spelling or grammatical errors.    
Analysis Question 1: (35%)  
Analysis Question 2: (35 %)  


```{r read data}

	setwd(data_dir)
	homes <- read.csv("train.csv", stringsAsFactors = FALSE)
	setwd(home_dir)

	names(homes) <- tolower(names(homes))
	
	for (i in 2:(length(homes)))
	{
		if (class(homes[,i]) == "character")
		{
			homes[,i] <- factor (homes[,i])
		}
	}
	

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	remove outliers ... more than 5 sigma from mean value
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
	
	lst <- length(homes) - 1	# sale price is (currently) last column
	
	for (i in 2 : lst)
	{
		if(class(homes[,i]) == "integer" || class(homes[,i]) == "numeric")
		{
			homes[,i][which(scale(homes[,i]) > 5)] <- NA
			homes[,i][which(scale(homes[,i]) < -5)] <- NA
		}
	}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	create a few new columns
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	dates <- paste(homes$yrsold, sprintf("%02d", homes$mosold), "01")
	homes$sale_date <- as.Date(dates, "%Y %m %d")
	
	homes$total_baths <- homes$fullbath + homes$halfbath / 2.0

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	scale each column independently
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

#	for (i in 2 : length(homes))
#	{
#		if(class(homes[,i]) == "integer" || class(homes[,i]) == "numeric")
#		{
#			homes[,i] <- scale(homes[,i])
#		}
#	}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	make some plots for numberic variables... linear, log_x, log_y, log_xy ...
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

#	pdf ("homes_train_plots.pdf", width = 10, height = 7)

	par (mfrow = c (2, 3))
	for (i in 2:(length(homes)))
	{
		if(class(homes[,i]) == "integer" || class(homes[,i]) == "numeric" || class(homes[,i]) == "matrix")
		{
			plot (homes[,i], main = (names(homes[i])))
			hist(homes[,i])
			plot(log(homes$saleprice)  ~ homes[,i])
		}
	}

	par (mfrow = c (1, 1))
	for (i in 2:(length(homes)))
	{
		if(class(homes[,i]) == "factor")
		{
			p <- ggplot(homes, aes(x = homes[,i], y = log(saleprice), fill = homes[,i])) + geom_boxplot()
			p + ggtitle(names(homes[i]))
			print(p)
		}
	}	
		
			plot(homes$saleprice ~ homes$sale_date)

#	dev.off()
	
	for (i in 2:(length(homes)))
	{
		if(class(homes[,i]) == "integer" || class(homes[,i]) == "numeric" || class(homes[,i]) == "matrix")
		{
			fit <- lm(homes$saleprice ~ homes[,i])
			
			print(sprintf(" ... %3d : %20s | r^2 = %8.3f | p-value = %12.4e",
						  i, names(homes[i]), summary(fit)$r.squared, summary(fit)$coefficients[,4][2] ))
		}
	}

	
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	Columns to remove - based on visual inspection
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
	
	homes_subset <- subset(homes, select = -c(
		id,
		mssubclass,
		street,
		alley,
		utilities,
		condition2,
		roofmatl,
		centralair,
		bsmtfinsf2,
		lowqualfinsf,
		bsmthalfbath,
		kitchenabvgr,
		x3ssnporch,
		screenporch,
		garagequal,
		garagecond,
		paveddrive,
		poolarea,
		poolqc,
		miscval,
		mosold,
		yrsold))

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	Impute NAs to functional value
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	for (i in 2 : (length(homes_subset)))
	{
		if(class(homes_subset[,i]) == "integer" || class(homes_subset[,i]) == "numeric" || class(homes_subset[,i]) == "matrix")
		{
			homes_subset[,i][is.na (homes_subset[,i])] <- min (homes_subset[,i], na.rm = TRUE)
		}
	}
	
	for (i in 2:(length(homes_subset)))
	{
		if(class(homes_subset[,i]) == "character")
		{
			homes_subset[,i][is.na (homes_subset[,i])] <- "None"
		}
	}

	homes_subset$log_saleprice <- log(homes_subset$saleprice)
	
	homes_subset <- subset(homes_subset, select = -c(saleprice, sale_date))
	
	sas_dir <- "~/sas/SASUniversityEdition/myfolders/"
	setwd(sas_dir)
	write.csv (homes_subset, file = "training_set_cleaned.csv", row.names = FALSE)

```
